---
title: "Cancer Classification"
author: "Komarathe"
date: "12/28/2021"
output: html_document
---

```{r load_packages}
# Load required packages
library(tidyverse)
library(caret)
library(class)
library(gmodels)
```

**Data Resource:**

- I gathered this data from [Kaggle](https://www.kaggle.com/ramadan16006/cancerclassification)
- 31 columns contain information regrading tumors size, symmetry, radius, perimeter, malignant or benign, etc.

```{r load_data}
# Load data 
tumor_data <- read.csv("cancer_classification.csv")
head(tumor_data)
```

## Data Exploration

```{r summary}
# Summary statistics
str(tumor_data)
summary(tumor_data)
```

```{r}
# Replace . in column names with _ 
colnames(tumor_data) <- gsub("\\.", "_", colnames(tumor_data))
```


**FINDINGS:**

- There are 569 observations and 31 variables
- All variables are numeric
- Our target variable is benign_0__mal_1, where value 1 represents malignant (cancer) and 0 represents benign
- All variables are measurements in different units
- There are no missing values in this data


```{r boxplots}
# Plot boxplots for all numeric variables
for (i in 1:30) {
  boxplot(tumor_data[,i], 
          main = paste("Box plot for", colnames(tumor_data)[i]), 
          ylab = colnames(tumor_data)[i])
}
```


```{r distribution}
# Plot distribution of all the independent variables
for(i in 1:30) {
  hist(tumor_data[,i], 
       main = colnames(tumor_data)[i], 
       xlab = colnames(tumor_data)[i], 
       col = 'yellow')
}

# Get frequency of values in traget variable
table(tumor_data$benign_0__mal_1)
```

### Scale the numeric variales for KNN Algorithm

- Since all the values in the numeric variables are measured in different units their ranges vary.
- For KNN algorithm we will scale these variables using min-max normalization.
```{r}
# Make a copy of original data
knn_data <- tumor_data

# Create a function for min-max normalization
min_max <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}

# Scale the data using min_max()
knn_norm_data <- as.data.frame(lapply(knn_data[1:30], min_max))

# Add benign_0__mal_1 column to knn_norm_data
knn_norm_data$benign_0__mal_1 <- knn_data$benign_0__mal_1
```


### Data preparation for KNN Algorithm
```{r}
# Divide data into training and testing data sets
knn_samples <- createDataPartition(knn_norm_data$benign_0__mal_1, p = 0.8, list = FALSE)
knn_train <- knn_norm_data[knn_samples,]
knn_test <- knn_norm_data[-knn_samples,]

# Store target variable of training and testing data sets into different vectors
knn_train_labels <- knn_train[, 31]
knn_test_labels <- knn_test[, 31]
```

### Train model on train data
For k, we will use square root of number of rows in train data which is 21
```{r}
# Use knn() from class package
knn_pred <- knn(train = knn_train, 
                test = knn_test, 
                cl = knn_train_labels, 
                k = 21)

```

### Evaluate KNN model
```{r}
# Get the confusion matrix of predicted and actual values
CrossTable(x = knn_test_labels, y = knn_pred, prop.chisq=FALSE)
```

**FINDINGS:**

- 49/113 were correctly identified as benign.
- 64/113 were correctly identified as malignant
- No false negatives and fals positives. The model correctly classified all 113 cases.


